---
title: "Forecast Swiss Retail Sales Index"
output: html_notebook
---

### Preliminaries 

```{r setup, warning=FALSE}
#install.packages("knitr")
library(knitr)
knitr::opts_knit$set(root.dir = "/Users/floriangoldinger/Desktop/Unilu/Forecasting in Economics and Business/Forecasting Project")
knitr::opts_chunk$set(message = FALSE,warning=FALSE)
```

```{r}
 rm(list = ls())
```

### Install and load packages 

```{r, comment=FALSE, warning=FALSE}
#install.packages("tseries")  
#install.packages("xts") 
#install.packages("forecast") 
#install.packages("tsbox")    
#install.packages("seasonal")
#install.packages("mFilter")
#install.packages("bruceR", dep=TRUE)
#install.packages("fanplot")

library(tseries)  
library(xts)
library(forecast) 
library(tsbox)    
library(seasonal)
library(mFilter)
library(dplyr)
library(readxl)
library(vars)
library(MASS)
library(bruceR)
library(fanplot)  
```

### Prepare and transform data 

Load data: 

```{r}
rawData <- read_excel("retail_sales_transp_March.xlsx")
```

Convert the data frame to a xts object and use the first column as date:

```{r}
data <- as.xts(x = rawData[, -1], order.by = as.Date(rawData$Date))
```

Take the first difference of the log to later work with the growth rates : 

```{r}
data$df_Retail_sales <- diff(log(data$Retail_sales), lag = 1, differences = 1)
data$df_Inflation <- diff(log(data$Inflation), lag = 1, differences = 1)
data$df12_Retail_sales <- diff(log(data$Retail_sales), lag = 12, differences = 1)
data$df_CHFEUR <- diff(log(data$CHFEUR), lag = 1, differences = 1)
data$df_Unemploy <- diff(log(data$Unemploy), lag = 1, differences = 1)
data$df_Brent_COIL <- diff(log(data$Brent_COIL), lag = 1, differences = 1)
data$df_Retail_EU <- diff(log(data$Retail_sales_EU), lag = 1, differences = 1)
```

Delete first row because of differencing 
```{r}
data <- tail(data, -1)
```

Visualize data 
```{r}
plot.xts(cbind(data$Retail_sales),  main = "Retail Sales Index (not seasonally adjusted)")
plot.xts(cbind(data$df_Retail_sales),  main = "Retail Sales Growth rate (not seasonally adjusted)")
plot.xts(cbind(data$Retail_sales_EU),  main = "Retail Sales EU")
plot.xts(cbind(data$df_Retail_sales),  main = "Retail Sales Growth rate (not seasonally adjusted)")
plot.xts(cbind(data$CHFEUR),  main = "CHF/EUR")
plot.xts(cbind(data$Inflation),  main = "Inflation")
plot.xts(cbind(data$Brent_COIL),  main = "Brent Crude Oil")
```
Seasonally adjust df Retail sales and df Retail Sales EU with monthly dummies.
```{r}
monthlyDummies <- forecast::seasonaldummy(ts_ts(data$df_Retail_sales)) 
modelretailseasonch <- lm(data$df_Retail_sales ~ monthlyDummies)
plot.xts(as.xts(modelretailseasonch$fitted.values),  main = "Swiss Retail Sales: fitted with seasonal dummy")
plot.xts(as.xts(modelretailseasonch$residuals),  main = "Swiss Retail Sales: deviations from seasonal dummy")
data$Retail_sales_cycle  <- modelretailseasonch$residuals

modelretailseasoneu <- lm(data$df_Retail_EU ~ monthlyDummies)
plot.xts(as.xts(modelretailseasoneu$fitted.values),  main = "EU Retail Sales: fitted with seasonal dummy")
plot.xts(as.xts(modelretailseasoneu$residuals),  main = "EU Retail Sales: deviations from seasonal dummy")
data$Retail_EU_cycle <- modelretailseasoneu$residuals
```
Create pandemic dummies to later exclude this period from analysis - suggesting a pandemic period from 01/01/2020 to 31/21/2021.

```{r}
pandemicDummies <- as.xts(ts(matrix(0, nrow = 320, ncol = 24), start = c(2000, 2), frequency = 12))
pandemicDummies["2020-01-01", 1] <- 1
pandemicDummies["2020-02-01", 2] <- 1
pandemicDummies["2020-03-01", 3] <- 1
pandemicDummies["2020-04-01", 4] <- 1
pandemicDummies["2020-05-01", 5] <- 1
pandemicDummies["2020-06-01", 6] <- 1
pandemicDummies["2020-07-01", 7] <- 1
pandemicDummies["2020-08-01", 8] <- 1
pandemicDummies["2020-09-01", 9] <- 1
pandemicDummies["2020-10-01", 10] <- 1
pandemicDummies["2020-11-01", 11] <- 1
pandemicDummies["2020-12-01", 12] <- 1
pandemicDummies["2021-01-01", 13] <- 1
pandemicDummies["2021-02-01", 14] <- 1
pandemicDummies["2021-03-01", 15] <- 1
pandemicDummies["2021-04-01", 16] <- 1
pandemicDummies["2021-05-01", 17] <- 1
pandemicDummies["2021-06-01", 18] <- 1
pandemicDummies["2021-07-01", 19] <- 1
pandemicDummies["2021-08-01", 20] <- 1
pandemicDummies["2021-09-01", 21] <- 1
pandemicDummies["2021-10-01", 22] <- 1
pandemicDummies["2021-11-01", 23] <- 1
pandemicDummies["2021-12-01", 24] <- 1
```

# Univariate model

ACF and PACF graphs to determine AC and MA lag 

```{r}
par(mfrow=c(1,2))
acf(data$Retail_sales_cycle,lag.max = 20, na.action = na.contiguous,  main = "Inclusive Pandemic")
acf(data["/2019-12-01"]$Retail_sales_cycle,lag.max = 20, na.action = na.contiguous,  main = "Exclusive Pandemic")

par(mfrow=c(1,2))
pacf(data$Retail_sales_cycle,lag.max = 20,na.action = na.contiguous,  main = "Inclusive Pandemic")
pacf(data["/2019-12-01"]$Retail_sales_cycle,lag.max = 20,na.action = na.contiguous,  main = "Exclusive Pandemic")
```
When excluding the pandemic period and thereafter: 
ACF/PACF suggests including 2 AC and 1 MA term 
Still some seasonality (non deterministic) - include a SARIMA model in a later stage? 

### Arima

Check with Arima (2,0,1) as suggested by AC and ACF including pandemic dummies - BIC of -1312.4

```{r}
forecast::Arima(data$Retail_sales_cycle,
          xreg = pandemicDummies["2000-02-01/2024-03-01"], 
          order = c(3L, 0L, 0L))
```
### Arima different lags

We check for the best AIC and MIC with different lags. Now a bigger model is chosen ARIMA(3,0,3) with the BIC information criterion. 
BIC -1337.453 but only minor improvement to more simple model. 

```{r}

nArOrder <- 4
nMaOrder <- 4 
matrixAIC <- matrix(data = NA,nArOrder+1,nMaOrder+1)
matrixBIC <- matrix(data = NA,nArOrder+1,nMaOrder+1)

for(iArOrder in 0:nArOrder){ 
  for(iMaOrder in 0:nMaOrder){
    temp = forecast::Arima(data$Retail_sales_cycle, 
                           order = c(iArOrder,0,iMaOrder), 
                           xreg = pandemicDummies["2000-01-01/2024-03-01"])
    print(iArOrder)
    print(iMaOrder)
    print(temp$bic)
    print(-2*temp$loglik + (iArOrder+iMaOrder)*log(length(data$Retail_sales_cycle)))
    matrixAIC[iArOrder+1,iMaOrder+1] = temp$aic
    matrixBIC[iArOrder+1,iMaOrder+1] = temp$bic
  }
}

which(matrixBIC==min(matrixBIC), arr.ind = TRUE)
which(matrixAIC==min(matrixAIC), arr.ind = TRUE)
```


### Auto Arima 

A lower BIC with the SARIMA model, which is not surprising as the AC and PAC still showed high seasonality with lag 12. 
A model SARIMA(0,0,1)(2,1,1)[12] is suggested by Auto Arima. We will use the SARIMA model also in the later steps to account for the seasonality as the BIC is lower: -1403.98. 

```{r}
modelretailsales <- forecast::auto.arima(ts_ts(data$df_Retail_sales),
                     ic="bic",
                     xreg = pandemicDummies["2000-02-01/2024-03-01"], 
                     max.p = 4,
                     max.q = 4,
                     trace = TRUE,
                     stepwise = FALSE)
```

We generate the forecast for the next 12 months based on the SARIMA(0,0,1)(1,0,1)[12] model.

```{r}
monthlyDummiesForecast <- forecast::seasonaldummy(ts_ts(data$Retail_sales), 12) 

forecastretail <- forecast::forecast(modelretailsales, 
                                     xreg = pandemicDummies["2024-04-01/2025-03-01"])
par(mfrow = c(1, 1))
plot(forecastretail, main = "Exclusive pandemic: SARIMA(0,0,1)(1,0,1)")
grid(nx = NULL, ny = NULL, lty = 3, col = "gray", lwd = 0.1)
```

# Multivariate model 

Explore the CCF Plot for different candidates for the exogenous variable.Cross-correlation only found for Retail EU (seasonally adjusted) - nothing for CHFEUR, Inflation, Brent_COIL or Unemployment.
We will assess if model can be improved by including Retail EU sales as exogenous variable. 

```{r}
ccf(as.ts(data$Retail_sales_cycle["/2019-12-31"]),as.ts(data$Retail_EU_cycle["/2019-12-31"]))
ccf(as.ts(data$Retail_sales_cycle["/2019-12-31"]),as.ts(data$df_CHFEUR["/2019-12-31"]))
ccf(as.ts(data$Retail_sales_cycle["/2019-12-31"]),as.ts(data$df_Inflation["/2019-12-31"]))
ccf(as.ts(data$Retail_sales_cycle["/2019-12-31"]),as.ts(data$df_Brent_COIL["/2019-12-31"]))
ccf(as.ts(data$Retail_sales_cycle["/2019-12-31"]),as.ts(data$df_Unemploy["/2019-12-31"]))
```

We double check the findings by modeling all the combinations and the related BIC values. Results:
Univariate 
-1403.98
Multivariate
Retail EU
-1429.2
Retail EU with 1 lag 
-1418.18
CHF/EUR
-1401.32
Retail EU AND CHF/EUR
-1425.24

Model with Retail EU (Lag 0) with lowest BIC and significant coefficient. 
Combined Retail EU and CHF/EUR with significant coefficient for CHF/EUR but no improvement in BIC. We decided to not include it in the forecast model because only low significance and we prefer a simple model. 

```{r}
forecast::auto.arima(ts_ts(data$df_Retail_sales),
                     ic="bic",
                     xreg = pandemicDummies["2000-02-01/2024-03-01"],
                     trace = TRUE,
                     stepwise = FALSE)

forecast::auto.arima(ts_ts(data$df_Retail_sales),
                     ic="bic",
                     xreg = cbind(pandemicDummies["2000-02-01/2024-03-01"], 
                                  data$df_Retail_EU),
                     trace = TRUE,
                     stepwise = FALSE)

forecast::auto.arima(ts_ts(data$df_Retail_sales),
                     ic="bic",
                     xreg = cbind(pandemicDummies["2000-02-01/2024-03-01"], 
                                  data$df_Retail_EU,
                                  lag(data$df_Retail_EU)
                     ),
                     trace = TRUE,
                     stepwise = FALSE)

forecast::auto.arima(ts_ts(data$df_Retail_sales),
                     ic="bic",
                     xreg = cbind(pandemicDummies["2000-02-01/2024-03-01"], 
                                  data$df_CHFEUR),
                     trace = TRUE,
                     stepwise = FALSE)

forecast::auto.arima(ts_ts(data$df_Retail_sales),
                     ic="bic",
                     xreg = cbind(pandemicDummies["2000-02-01/2024-03-01"], 
                                  data$df_Retail_EU,
                                  data$df_CHFEUR),
                     trace = TRUE,
                     stepwise = FALSE)

```

## Model with Exogenous variable Retail EU 

Model the multivariate model with Auto Arima to account for the seosonality. 

AR/MA terms for the evaluation (see chapter Evaluation):

Retail_EU_cycle: ARIMA(2,0,3)
ARIMA(0,0,1)(0,1,1)[12]
CHFEUR: ARIMA(0,0,1)
Retail_EU_cycle AND CHFEUR: ARIMA(0,0,1)
ARIMA(1,0,1)(2,1,0)[12]

```{r}
modelretailsales <- forecast::auto.arima(ts_ts(data$df_Retail_sales),
                     ic="bic",
                     xreg = cbind(pandemicDummies["2000-02-01/2024-03-01"], 
                                  data$df_Retail_EU
                                  ),
                     trace = TRUE,
                     stepwise = FALSE)

# Check AR/MA terms for later use in the evaluation part Retail EU 
forecast::auto.arima(ts_ts(data$df_Retail_EU),
                     ic="bic",
                     xreg = pandemicDummies["2000-02-01/2024-03-01"], 
                     trace = TRUE,
                     stepwise = FALSE)


# Check AR/MA terms for later use in the evaluation part CHF/EUR 
forecast::auto.arima(data$df_CHFEUR,
                     ic="bic",
                     xreg = pandemicDummies["2000-02-01/2024-03-01"],
                     trace = TRUE,
                     stepwise = FALSE)

# Check AR/MA terms for later use in the evaluation part Retail EU AND CHF/EUR 
forecast::auto.arima(ts_ts(data$df_Retail_sales),
                     ic="bic",
                     xreg = cbind(pandemicDummies["2000-02-01/2024-03-01"], 
                                  data$df_Retail_EU,
                                  data$df_CHFEUR),
                     trace = TRUE,
                     stepwise = FALSE)

```

Model and plot forecast for Retail EU 

```{r}
modelretaileu <- forecast::auto.arima(ts_ts(data$df_Retail_EU), 
                                         xreg = pandemicDummies["2000-02-01/2024-03-01"], 
                                         ic = "bic",
                                         trace = TRUE,
                                         stepwise = FALSE
                                         )

modelretaileu$bic  

forecasteu <- forecast::forecast(modelretaileu, 
                                 xreg = pandemicDummies["2024-04-01/2025-03-01"])

plot(forecasteu)
```

Forecast Swiss Retail with exogneous variable Euro Retail 

```{r}
forecastretail <- forecast::forecast(modelretailsales, 
                                     xreg = cbind(pandemicDummies["2024-04-01/2025-03-01"],
                                     forecasteu$mean))

plot(forecastretail)

```

Reverse the differencing and log to get a point forecast of the index level. 
Add on to the last observed value and convert back to the original scale. 

```{r}
cumulative_forecasted_diffs <- cumsum(forecastretail$mean[1:12])

forecasted_series_final <- cumulative_forecasted_diffs + log(tail(data$Retail_sales, 1)[[1]])

forecasted_series_original_scale <- exp(forecasted_series_final)

print(forecasted_series_original_scale)
```

# Evaluation 

Now for evaluation and for simplicity only use the data before the pandemic and non seasonal model. 
6 models are evaluated and the AR/MA terms that have been determined in the previous analysis. 
1) Univariate m/m
2) Univariate y/y
3) Multivariate with Retail EU 
4) Multivariate with Retail EU AND CHF/EUR
5) Naive "no change"
6) Average of the above 5 models 

```{r}
dataeval <- data[as.Date(index(data)) <= as.Date("2020-01-01"), ]

tEvaluationPeriods <- 70
tPeriods = length(dataeval[,"Retail_sales"])
forecastHorizon = 12
nModels = 6 # we compare these models: univariate m/m, y/y, multivariate with EU Retail, CHFEUR, naive, average
forecastRolling <- matrix(,tEvaluationPeriods,nModels)
for (tauEvaluationPeriod in 1:tEvaluationPeriods){
  
  lastPeriod <- tPeriods-tEvaluationPeriods+tauEvaluationPeriod-forecastHorizon 
  
  # Univariate model, m/m 
  modelRetail_temp <- forecast::Arima(ts_ts(dataeval$df_Retail_sales[1:lastPeriod]),
                                        order = c(0, 0, 1), seasonal=c(2, 1, 1))
  forecastRetail_temp <- forecast::forecast(modelRetail_temp, h = forecastHorizon)
  forecastRolling[tauEvaluationPeriod,1] <- sum(forecastRetail_temp$mean[1:forecastHorizon]) # y/y forecast is sum of m/m (with logs)

  # Univariate model, y/y
  modelRetail_temp = forecast::Arima(dataeval$df12_Retail_sales[12:lastPeriod],
                                        order = c(1, 0, 1))
  forecastRetail_temp = forecast::forecast(modelRetail_temp, h = forecastHorizon)
  forecastRolling[tauEvaluationPeriod,2] <- forecastRetail_temp$mean[forecastHorizon]

  # Multivariate model
  modelRetailExo_temp = forecast::Arima(ts_ts(dataeval$df_Retail_sales[1:lastPeriod]),
                                        order = c(0, 0, 1), seasonal=c(2, 1, 0),
                                        xreg = dataeval$df_Retail_EU[1:lastPeriod])
  modelRetaileu_temp <- forecast::Arima(ts_ts(dataeval$df_Retail_EU[1:lastPeriod]), 
                                        order = c(0, 0, 1), seasonal=c(0, 1, 1))
  forecasteu_temp <- forecast::forecast(modelRetaileu_temp, 
                                        h = forecastHorizon)  
  forecastRetail_temp <- forecast::forecast(modelRetailExo_temp,
                                            xreg = forecasteu_temp$mean,
                                            h = forecastHorizon)  
  forecastRolling[tauEvaluationPeriod,3] <- sum(forecastRetail_temp$mean[1:forecastHorizon])    
  
  # Multivariate with Retail EU and CHFEUR 
  
  modelRetailExo_temp = forecast::Arima(ts_ts(dataeval$df_Retail_sales[1:lastPeriod]),
                                        order = c(1, 0, 1), seasonal=c(2, 1, 0),
                                        xreg = cbind(dataeval$df_Retail_EU[1:lastPeriod],
                                                     dataeval$df_CHFEUR[1:lastPeriod]))
  modelRetaileu_temp <- forecast::Arima(ts_ts(dataeval$df_Retail_sales[1:lastPeriod]), 
                                        order = c(0, 0, 1), seasonal=c(0, 1, 1))
  modelCHFEUR_temp <- forecast::Arima(ts_ts(dataeval$df_CHFEUR[1:lastPeriod]), 
                                      order = c(0, 0, 1))
  forecasteu_temp <- forecast::forecast(modelRetaileu_temp, 
                                        h = forecastHorizon)  
  forecastCHFEUR_temp <- forecast::forecast(modelCHFEUR_temp, h = forecastHorizon)
  forecastRetail_temp <- forecast::forecast(modelRetailExo_temp,
                                            xreg = cbind(forecasteu_temp$mean, 
                                                         forecastCHFEUR_temp$mean),
                                            h = forecastHorizon)  
  forecastRolling[tauEvaluationPeriod,4] <- sum(forecastRetail_temp$mean[1:forecastHorizon]) 
  
  # "Naive", no change forecast
  
  forecastRolling[tauEvaluationPeriod,5] <- dataeval$df12_Retail_sales[lastPeriod]
  
  # average - just take the average of different models 
  forecastRolling[tauEvaluationPeriod,6] <- mean(forecastRolling[tauEvaluationPeriod,1:5])

}

matplot(1:70, cbind(forecastRolling,dataeval$df12_Retail_sales[171:240])) 
```

Now we evaluate the different models according to their performance using the quadratic loss and the absolute loss. 
Best model is the average of the models, followed by the univarite model. - worst is the naive model. 

```{r}
## Calculate the forecast errors 

actualData <- dataeval$df12_Retail_sales[(length(dataeval$df12_Retail_sales)-tEvaluationPeriods+1):length(dataeval$df12_Retail_sales)]
actualDataMatrix <- kronecker(matrix(1,1,6),actualData)
forecastErrors <- actualDataMatrix - forecastRolling

## Calculate the loss function (quadratic error and absolute error)

quadraticLoss <- forecastErrors^2 
absoluteLoss <- abs(forecastErrors)

## Calculate the mean loss for each model 

colMeans(quadraticLoss[1:70,])*100
colMeans(absoluteLoss[1:70,])*100

```

Check for significance difference between the multivariate with Retail EU and the univariate. The difference is significant with the corrected t-value.
It is significant. We might favor the univariate model. 

```{r}
regressionDMW <- lm((quadraticLoss[1:70,1]-quadraticLoss[1:70,3]) ~ 1)
summary(regressionDMW)

## t-value corrected for autocorrelation 

tHAC <- regressionDMW$coefficients/(NeweyWest(regressionDMW,lag = 12))^0.5
print(tHAC)
```

# Density forecast multivariate 

Model again with the whole timeframe but excluding the pandemic period. 

```{r}
modelretailsales <- forecast::auto.arima(ts_ts(data$df_Retail_sales),
                                         ic="bic",
                                         xreg = cbind(pandemicDummies["2000-02-01/2024-03-01"], 
                                                      data$df_Retail_EU
                                         ),
                                         max.d = 0, 
                                         trace = TRUE,
                                         stepwise = FALSE)

modelretaileu <- forecast::auto.arima(ts_ts(data$df_Retail_EU), 
                                      xreg = pandemicDummies["2000-02-01/2024-03-01"], 
                                      ic = "bic",
                                      trace = TRUE,
                                      stepwise = FALSE
                                      )
```


Density forecast with random draw from normal distribution of error term 

```{r}
tHorizons <- 12
nDraws <- 100
forecastRetailMatrix <- rbind(kronecker(matrix(1,1,nDraws),data.matrix(data$df_Retail_sales)),matrix(NA,tHorizons,nDraws))
forecastEUMatrix <- rbind(kronecker(matrix(1,1,nDraws),data.matrix(data$df_Retail_EU)),matrix(NA,tHorizons,nDraws))

pandemicDummiesMatrix <- kronecker(matrix(1,1),pandemicDummies)

T <- length(data$df_Retail_sales)
for (tauHorizon in 1:tHorizons) {
  
  for (iDraw in 1:nDraws) {

    modelEU_temp <- Arima(forecastEUMatrix[1:T+tauHorizon-1,iDraw],model = modelretaileu, xreg = pandemicDummiesMatrix[1:T+tauHorizon-1,]) 
    
    meanForecastEU <- forecast::forecast(modelEU_temp,h=1,xreg = matrix(pandemicDummiesMatrix[(T+tauHorizon),],1,24))$mean 
    
    forecastEUMatrix[T+tauHorizon,iDraw] <- meanForecastEU + rnorm(1, mean=0, sd=modelretaileu$sigma2^0.5)
    
    
    modelRetail_temp <- forecast::Arima(forecastRetailMatrix[1:T+tauHorizon-1,iDraw],model = modelretailsales, xreg = cbind(forecastEUMatrix[1:T+tauHorizon-1,iDraw],pandemicDummiesMatrix[1:T+tauHorizon-1,]))
    
    meanForecastRetail <- forecast::forecast(modelRetail_temp,h=1,xreg = cbind(forecastEUMatrix[T+tauHorizon,iDraw],matrix(pandemicDummiesMatrix[(T+tauHorizon),],1,24)))$mean
    forecastRetailMatrix[T+tauHorizon,iDraw] <- meanForecastRetail + rnorm(1, mean=0, sd=modelretailsales$sigma2^0.5)
    
  }
} 

plot(as.ts(data.matrix(data$df_Retail_sales)), type = "l", lwd = 2, las = 1, ylab = "",xlim = c(1,T+tHorizons))
fan(data = t(forecastRetailMatrix[(T+1):(T+tHorizons),]), type = "interval", start = (T+1))
```

Density forecast with ootstrapping 

```{r}
forecastRetailMatrixboot <- rbind(kronecker(matrix(1,1,nDraws),data.matrix(data$Retail_sales_cycle)),matrix(NA,tHorizons,nDraws))
forecastEUMatrixboot <- rbind(kronecker(matrix(1,1,nDraws),data.matrix(data$df_Retail_EU)),matrix(NA,tHorizons,nDraws))

modelretaileu$residuals <- modelretaileu$residuals - mean(modelretaileu$residuals)
modelretailsales$residuals <- modelretailsales$residuals - mean(modelretailsales$residuals)

T <- length(data$Retail_sales_cycle)
for (tauHorizon in 1:tHorizons) {
  
  for (iDraw in 1:nDraws) {
    
    modelEU_temp <- Arima(forecastEUMatrix[1:T+tauHorizon-1,iDraw],model = modelretaileu, xreg = pandemicDummiesMatrix[1:T+tauHorizon-1,]) 
    
    meanForecastEU <- forecast::forecast(modelEU_temp,h=1,xreg = matrix(pandemicDummiesMatrix[(T+tauHorizon),],1,24))$mean
    
    forecastEUMatrixboot[T+tauHorizon,iDraw] <- meanForecastEU + modelretaileu$residuals[ceiling(runif(1,0,length(modelretaileu$residuals)))]
    
    
    modelRetail_temp <- forecast::Arima(forecastRetailMatrixboot[1:T+tauHorizon-1,iDraw],model = modelretailsales,xreg =     cbind(forecastEUMatrixboot[1:T+tauHorizon-1,iDraw],pandemicDummiesMatrix[1:T+tauHorizon-1,]))
    
    meanForecastRetail <- forecast::forecast(modelRetail_temp,h=1,xreg = cbind(forecastEUMatrixboot[T+tauHorizon,iDraw],matrix(pandemicDummiesMatrix[(T+tauHorizon),],1,24)))$mean
    forecastRetailMatrixboot[T+tauHorizon,iDraw] <- meanForecastRetail + modelretailsales$residuals[ceiling(runif(1,0,length(modelretailsales$residuals)))]
  }
  
} 

plot(as.ts(data.matrix(data$df_Retail_sales)), type = "l", lwd = 2, las = 1, ylab = "",xlim = c(1,T+tHorizons))
fan(data = t(forecastRetailMatrixboot[(T+1):(T+tHorizons),]), type = "interval", start = T+1)
```







